---
title: "worldpa: an interface to the World Database on Protected Areas (WDPA)"
author: "Nicolas CASAJUS < nicolas.casajus@fondationbiodiversite.fr >"
date: "`r Sys.Date()`"
lang: english
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{An introduction of the package worldpa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Introduction

This R package is an interface to the World Database on Protected Areas (WDPA)
hosted in the Protected planet website: [https://www.protectedplanet.net](https://www.protectedplanet.net).
It allows user to download spatial shapefiles of protected areas (PA) for world countries.


# Installation

To install this package from [GitHub](https://www.github.com/frbcesab/worldpa),
first install the package [`devtools`](http://cran.r-project.org/web/packages/devtools/index.html) from the CRAN (if not already installed).

```{r echo=TRUE, eval=FALSE}
# Install the devtools package
install.packages("devtools", dependencies = TRUE)

# Install the worldpa package
devtools::install_github("frbcesab/worldpa", build_vignettes = TRUE)

# Load the worldpa package
library(worldpa)
```

```{r echo=FALSE, eval=TRUE}
library(worldpa)
```


# Getting started

This package contains four main functions:

- [`get_token()`](https://frbcesab.github.io/worldpa/reference/get_token.html): checks if your WDPA token is correctly stored;
- [`get_countries()`](https://frbcesab.github.io/worldpa/reference/get_countries.html): returns the world countries names according to the UNEP classification;
- [`get_wdpa()`](https://frbcesab.github.io/worldpa/reference/get_wdpa.html): downloads world protected areas shapefiles.

Let's explore the function [`get_token()`](https://frbcesab.github.io/worldpa/reference/get_token.html).

```{r echo=TRUE, eval=FALSE}
# Checks if your WDPA token is correctly stored
get_token()
```

```{r echo=FALSE, eval=TRUE}
Sys.getenv("WDPA_FAKE")
```
