---
title: "worldpa: An Interface to the World Database on Protected Areas (WDPA)"
author: "Nicolas CASAJUS"
date: "`r Sys.Date()`"
lang: english
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{An introduction of the package worldpa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Overview

<hr />



The R package `{worldpa}` is an interface to the World Database on Protected Areas (WDPA)
hosted on the Protected Planet website: [https://www.protectedplanet.net](https://www.protectedplanet.net).
It allows users to download spatial shapefiles of terrestrial and marine protected areas (PA) for world countries using the WDPA API ([https://api.protectedplanet.net](https://api.protectedplanet.net)).

This package contains three main functions:

- `get_token`: this function checks if your WDPA personal access token (PAT) is correctly stored;
- `get_countries`: this function gets world countries information (ISO-3 code);
- `get_wdpa`: this function downloads world protected areas shapefiles for one specified country.

<br />

This package is not yet hosted on the CRAN but you can install it from 
[GitHub](https://www.github.com):

```{r echo=TRUE, eval=FALSE}
# install.packages("remotes")
remotes::install_github("FRBCesab/worldpa")
```



<br />

# Managing WDPA API Token

<hr />



This package uses the WDPA API to access data on the World protected areas. You must
first have obtained a Personal API PAT by filling in the form available at:
[https://api.protectedplanet.net/request](https://api.protectedplanet.net/request).
Once you have obtained this token, you must store it on your computer.
As recommended by Hadley Wickham in his 
[Managing secrets vignette](https://cran.r-project.org/web/packages/httr/vignettes/secrets.html),
you can store your API PAT as an R Environment variable (a name-value pairs)
listed in the `.Renviron` file in your home directory.

Use the function `usethis::edit_r_environ` to open this file and add this line 
(by replacing `z999zzz...` with your Personal WDPA API Token):

```
WDPA_KEY='z999zzz9zzz999999z9z99zz999zz999'
```

Save the file and relaunch R.

Now, let's check if your Personal WDPA API Token is correctly stored in your 
`.Renviron` file by using the function `get_token`:

```{r echo=TRUE, eval=FALSE}
## Check if User's Token is correctly stored ----
wdpa::get_token()

## [1] "z999zzz9zzz999999z9z99zz999zz999"
```



<br />

# Getting Countries Information

<hr />



The internal dataset `wdpa_countries` contains information on world countries 
(country and UNEP region names, ISO code and number of protected areas).


```{r echo=TRUE, eval=FALSE}
data("wdpa_countries", package = "worldpa")

##    region_name region_iso2             country_name country_iso3 pas_count
## 1       Africa          AF                  Algeria          DZA        78
## 2       Africa          AF                   Angola          AGO        14
## 3       Africa          AF                    Benin          BEN        64
## 4       Africa          AF                 Botswana          BWA        22
## 5       Africa          AF             Burkina Faso          BFA       112
## 6       Africa          AF                  Burundi          BDI        21
## 7       Africa          AF               Cabo Verde          CPV         7
## 8       Africa          AF                 Cameroon          CMR        49
## 9       Africa          AF Central African Republic          CAF        38
## 10      Africa          AF                     Chad          TCD        23
```

<br />

As the WDPA is updated every month, you can get an updated version of this 
dataset by using the function `get_countries`:


```{r echo=TRUE, eval=FALSE}
## Get UNEP-WCMC World Countries Information ----
wdpa::get_countries()
```



<br />

# Downloading Protected Areas

<hr />



Now, let's use the function `get_wdpa` to download protected areas shapefiles 
for Georgia.


```{r echo=TRUE, eval=FALSE}
## Get Georgia ISO-3 Code ----
wdpa_countries[grep("Georgia", wdpa_countries$"country_name"), 3:4]

##                                     country_name country_iso3
## 133                                      Georgia          GEO
## 235 South Georgia and the South Sandwich Islands          SGS

# Download Georgia PA ----
pa_sf <- get_wdpa(isocode = "GEO")
```

The returned object is a simple feature **sf object** of `MULTIPOLYGONS` as 
defined in the package `{sf}`. The function `get_wdpa` has also exported the 
protected areas as an ESRI Shapefile in the working directory.

<br />

If you have any question, please open an issue at: 
[https://github.com/frbcesab/worldpa/issues](https://github.com/frbcesab/worldpa/issues)
