---
title: "worldpa: an interface to the World Database on Protected Areas (WDPA)"
author: "Nicolas CASAJUS < nicolas.casajus@fondationbiodiversite.fr >"
date: "`r Sys.Date()`"
lang: english
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{An introduction of the package worldpa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Introduction

This R package is an interface to the World Database on Protected Areas (WDPA)
hosted in the Protected planet website: [https://www.protectedplanet.net](https://www.protectedplanet.net).
It allows user to download spatial shapefiles of protected areas (PA) at three different levels: country, region or the world.


# Installation

To install this package from [GitHub](https://www.github.com/frbcesab/worldpa),
first install the package [`devtools`](http://cran.r-project.org/web/packages/devtools/index.html) from the CRAN (if not already installed).

```{r echo=TRUE, eval=FALSE}
# Install the devtools package   ----------
install.packages("devtools", dependencies = TRUE)

# Install the worldpa package    ----------
devtools::install_github("frbcesab/worldpa", build_vignettes = TRUE)

# Load the worldpa package       ----------
library(worldpa)
```

```{r echo=FALSE, eval=TRUE}
library(worldpa)
```


# Getting started

...
