---
title: "worldpa: An interface to the World Database on Protected Areas (WDPA)"
author: "Nicolas CASAJUS < nicolas.casajus@fondationbiodiversite.fr >"
date: "`r Sys.Date()`"
lang: english
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{An introduction of the package worldpa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


# Introduction

<hr />

This R package is an interface to the World Database on Protected Areas (WDPA)
hosted on the Protected planet website: [https://www.protectedplanet.net](https://www.protectedplanet.net).
It allows user to download spatial shapefiles of protected areas (PA) for world countries using the WDPA API ([https://api.protectedplanet.net](https://api.protectedplanet.net)).


This package contains three main functions:

- [`get_token()`](https://frbcesab.github.io/worldpa/reference/get_token.html): check if your WDPA token is correctly stored;
- [`get_countries()`](https://frbcesab.github.io/worldpa/reference/get_countries.html): get world countries informations;
- [`get_wdpa()`](https://frbcesab.github.io/worldpa/reference/get_wdpa.html): download world protected areas shapefiles for one specified country.

<br />

To install this package from [GitHub](https://www.github.com/frbcesab/worldpa),
first install the package [`devtools`](http://cran.r-project.org/web/packages/devtools/index.html) from the CRAN (if not already installed).

```{r echo=TRUE, eval=FALSE}
# Install the {devtools} package
install.packages("devtools", dependencies = TRUE)

# Install the {worldpa} package
devtools::install_github("frbcesab/worldpa")

# Load the {worldpa} package
library(worldpa)

# Print the package content
ls("package:worldpa")

# Browse the package help
browseURL("https://frbcesab.github.io/worldpa/reference/")

# Read this vignette
browseURL("https://frbcesab.github.io/worldpa/articles/worldpa.html")
```

```{r echo=FALSE, eval=TRUE}
library(worldpa)
```

<br />

# Managing WDPA API token

<hr />


This package uses the WDPA API to access data on world protected areas. You must
first have obtained a Personal API Token by filling in the form available at:
[https://api.protectedplanet.net/request](https://api.protectedplanet.net/request).
Once you have obtained this token, you must store it on your computer.
As recommended by Hadley Wickham in it [Managing secrets vignette](https://cran.r-project.org/web/packages/httr/vignettes/secrets.html),
you have to stored your API Token as an R Environment variable (a name-value pairs)
listed in the `.Renviron` file in your home directory.

Use the function `usethis::edit_r_environ()` to open this file and add this line (by replacing `z999zzz...` with your Personal WDPA API Token):

```
WDPA_KEY=z999zzz9zzz999999z9z99zz999zz999
```

Save the file and relaunch R.

Now, let's check if your Personal WDPA API Token is correctly stored in your `.Renviron` file by using the function [`get_token()`](https://frbcesab.github.io/worldpa/reference/get_token.html).

```{r echo=TRUE, eval=FALSE}
# Checks if your WDPA token is correctly stored
get_token()
```

```{r echo=TRUE, eval=FALSE}
## [1] "z999zzz9zzz999999z9z99zz999zz999"
```

<br />

# Getting countries metadata

<hr />

The internal dataset `wdpa_countries` contains different informations on world countries (country and UNEP region names, ISO code and number of protected areas).


```{r echo=TRUE, eval=TRUE}
# Load internal dataset
data(wdpa_countries)
```

```{r echo=FALSE, eval=TRUE}
# Print first rows of the dataset
head(wdpa_countries, 10)
```

<br />

As the WDPA is updated every month, you can get an updated version of this dataset by using the function [`get_countries()`](https://frbcesab.github.io/worldpa/reference/get_countries.html):

```{r echo=TRUE, eval=FALSE}
# Get UNEP world countries informations
get_countries()
```


<br />

# Getting protected areas shapefiles

<hr />

```{r echo=TRUE, eval=TRUE}

wdpa_countries[grep("Georgia", wdpa_countries[ , "country_name"]), ]

pa_sf <- get_wdpa(isocode = "GEO")

print(pa_sf)

dir()

plot(pa_sf)
```
