---
title: "worldpa: An interface to the World Database on Protected Areas (WDPA)"
author: "Nicolas Casajus"
date: "`r Sys.Date()`"
lang: english
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{An introduction of the package worldpa}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



# Introduction

This R package is an interface to the World Database on Protected Areas (WDPA)
hosted in the Protected planet website: [https://www.protectedplanet.net](https://www.protectedplanet.net).
It allows user to download spatial shapefiles of protected areas (PA) at three different levels: country, region or the world.


# Installation

To install this package from [GitHub](https://www.github.com/frbcesab/worldpa),
first install the package [`devtools`](http://cran.r-project.org/web/packages/devtools/index.html) from the CRAN (if not already installed).

```{r echo=TRUE, eval=FALSE}
# Install the devtools package   ----------
install.packages("devtools", dependencies = TRUE)

# Install the worldpa package    ----------
devtools::install_github("frbcesab/worldpa", build_vignettes = TRUE)

# Load the worldpa package       ----------
library(worldpa)
```

```{r echo=FALSE, eval=TRUE}
# Get the world regions names    ----------
library(worldpa)
```


# Getting started

This package contains four main functions:

- `get_regions()`: returns the world regions names according to the UNEP classification
- `get_countries()`: returns the world countries names according to the UNEP classification
- `get_isocodes()`: returns the ISO3 code of all countries of the world
- `get_wdpa()`: downloads world protected areas shapefiles (core function)

Let's explore the function `get_regions()`.

```{r echo=TRUE, eval=TRUE}
# Get the world regions names    ----------
get_regions()
```

This function returns the correct spelling of world regions, useful to set the parameter `regions` of the function `get_wdpa()`.

Now, let's take a look at the function `get_countries()`.

```{r echo=TRUE, eval=TRUE}
# Get the world countries names  ----------
countries <- get_countries()

# Get the number of countries    ----------
length(countries)

# Print the first ten countries  ----------
countries[1:10]
```

This function returns the correct spelling of world countries, useful to set the parameter `countries` of the function `get_wdpa()`.

The function `get_isocodes()` returns the ISO3 code (three-letters acronym) of all countries of the world.

```{r echo=TRUE, eval=TRUE}
# Get the countries ISO3 codes   ----------
iso3 <- get_isocodes()

# Print the first twenty rows    ----------
head(iso3, 20)
```

Finally, the use of the function `get_wdpa()` is quite simple:

```{r echo=TRUE, eval=FALSE}
# Download PA for one country    ----------
get_wdpa(countries = "Spain")

# Download PA for two countries  ----------
get_wdpa(countries = c("Portugal", "Spain"))

# Download PA for one region     ----------
get_wdpa(regions = "Africa")

# Download PA for two regions    ----------
get_wdpa(regions = c("Europe", "Africa"))

# Download PA for the world      ----------
get_wdpa() # Not recommended (too big)
```

The argument `path` can be set to change the directory where shapefiles will be written. The default is the current directory.
Let's say we want to save shapefiles in the directory **wdpa** located on the **Desktop**:

```{r echo=TRUE, eval=FALSE}
get_wdpa(countries = "Spain", path = "/Users/username/Desktop/wdpa")   # Unix
get_wdpa(countries = "Spain", path = "C:/Users/username/Desktop/wdpa") # Windows
```

**Note:** the function `get_wdpa()` does not return any object on R: shapefiles are directly written on the hard drive.


# Study case

```{r echo=TRUE, eval=FALSE}
get_wdpa(regions = "Africa", path = "Desktop/WDPA")
```
